---
title: "Jointly optimize nu and delta"
Author: "Dat Do"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
---

Consider only model 3 and 5 from Markdown "Summary of learning nu in all considered models" and now we vary both $\nu$ and $\delta$.


(Model 3) $\mathrm{F}(R_0 | \frac{\nu \Psi'}{N}, N, \nu + 2)$

(Model 5) $\mathcal{IW}(R_0 | \frac{\nu \Psi'}{N}, \nu + p + 1)$


where $\Psi' = \dfrac{\nu R' + \delta \text{diag}(R_0)}{\nu + \delta}$.

## 1. Data preparation and log-likelihood functions
First, load the data:

```{r}
seed = 10  ## change this to see different experiment
set.seed(seed)
library(susieR)
library(Matrix)

data(N3finemapping)
attach(N3finemapping)
X0 = N3finemapping$X
## getting covariance matrix from the whole sample
## and examine the eigendecomposition to estimate numerical rank
R = cov(X0)
eig <- eigen(R)
plot(eig$values,
     main = "Eigenvalues of covariance matrix calculated using all samples",
     ylab = "Value",
     xlab = "Eigenvalue index")

n0 = dim(X0)[1]
p0 = dim(X0)[2]
snp_total = p0
```

We proceed to split the data into half and look at the heatmap of the covariance matrices of two sub-samples.

```{r}
#### randomly split the data into half
#### randomly select p consecutive SNPs where p < n so IW is proper
p = 100
# Start from a random point on the genome
indx_start = sample(1: (snp_total - p), 1)
X = X0[, indx_start:(indx_start + p -1)]
# View(cor(X)[1:10, 1:10])
## sub-sample into two
out_sample_size = n0 / 2
out_sample = sample(1:n0, out_sample_size)
X_out = X[out_sample, ]
X_in = X[setdiff(1:n0, out_sample), ]

rm_p = c(which(diag(cov(X_in))==0), which(diag(cov(X_out))==0))
indx_p = setdiff(1:p, rm_p)
X_in = X_in[, indx_p]
X_out = X_out[, indx_p]
## out-sample LD matrix
p = length(indx_p)
Rp = cov(X_out)
R0 = cov(X_in)
library(ggplot2)
library(reshape2)
df1 <- melt(R0)
df2 <- melt(Rp)
N_in = nrow(X_in)
N_out = nrow(X_out)
p1 <- ggplot(df1, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  coord_fixed() +
  ggtitle(sprintf("In-sample Cov, %d samples", nrow(X_in)))
p2 <- ggplot(df2, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  coord_fixed() +
  ggtitle(sprintf("Out-of-sample Cov, %d samples", nrow(X_out)))
library(gridExtra)
grid.arrange(p1, p2, ncol = 2)
```

Log-likelihood functions: 

```{r}
## Matrix F-distribution likelihood
#### log F(R0 | nu * Rp / N, N, nu + 2)
log_multigamma_vec <- function(a, p) {
  # vectorized multivariate gamma
  shifts <- (1 - seq_len(p)) / 2
  # sum over j, but broadcasting a over j
  (p*(p-1)/4)*log(pi) +
    rowSums(lgamma(outer(a, shifts, "+")))
}


log_F <- function(R0, Rp, N, nu_vec) {
  p <- nrow(R0)
  jitter = 1e-8
  R0 = R0 + jitter * diag(p)
  Rp = Rp + jitter * diag(p)
  # Precompute expensive shared quantities
  logdet_nu_Rp_over_N <- (determinant(Rp, logarithm = TRUE)$modulus 
                          + p * log(nu_vec)
                          - p * log(N))
  logdetR0   <- determinant(R0, logarithm = TRUE)$modulus
  
  # lambda_vec <- eigen(solve(Rp, R0))$values
  # lambda_over_nu = tcrossprod(lambda_vec, N / nu_vec)
  # logdet_I_plus_RR <- colSums(log(1 + lambda_over_nu))
  # llhs = (log_multigamma_vec((N + nu_vec + p + 1) / 2, p)
  #         - log_multigamma_vec(N / 2, p)
  #         - log_multigamma_vec((nu_vec + p + 1) / 2, p)
  #         - .5 * N * logdet_nu_Rp_over_N
  #         + .5 * (N - p - 1) * logdetR0
  #         - .5 * (N + nu_vec + p + 1) * logdet_I_plus_RR)
  
  logdet_Rplus_Rp = rep(0, length(nu_vec))
  for (idx in 1:length(nu_vec)){
    nu = nu_vec[idx]
    logdet_Rplus_Rp[idx] <- determinant(R0 + nu * Rp / N, logarithm = TRUE)$modulus
  }
  
  llhs = (log_multigamma_vec((N + nu_vec + p + 1) / 2, p)
          - log_multigamma_vec(N / 2, p)
          - log_multigamma_vec((nu_vec + p + 1) / 2, p)
          + .5 * (nu_vec + p + 1) * logdet_nu_Rp_over_N
          + .5 * (N - p - 1) * logdetR0
          - .5 * (N + nu_vec + p + 1) * logdet_Rplus_Rp)
  
  as.numeric(llhs)
}

log_iw <- function(R0, Rp, nu_vec) {
  p <- nrow(R0)
  jitter = 1e-12
  R0 = R0 + jitter * diag(p)
  Rp = Rp + jitter * diag(p)
  # Precompute expensive shared quantities
  logdet_nu_Rp <- determinant(Rp, logarithm = TRUE)$modulus + p * log(nu_vec)
  logdetR0   <- determinant(R0,   logarithm = TRUE)$modulus
  tr_term   <- nu_vec * sum(t(Rp) * solve(R0))
  llhs = (.5 * (nu_vec + p + 1) * logdet_nu_Rp
          - .5 * (nu_vec + p + 1) * p * log(2)
          - log_multigamma_vec((nu_vec + p + 1) / 2, p)
          - .5 * (nu_vec + 2 * (p + 1)) * logdetR0
          - .5 * tr_term)
  as.numeric(llhs)
}


```

## 2. Comparing $\nu$ learned in the F model with various delta

Sorry In this simulation $\Psi' = \dfrac{N' R' + \text{diag}(R_0)}{N' + \delta}$

This is not exactly setting $\Psi' = \dfrac{\nu R' + \text{diag}(R_0)}{\nu' + \delta}$ as we want. But I found it still gives some more intuition.

```{r}
N = nrow(X_in)
Np = nrow(X_out)
nu_vec = c(1:100) 
delta = 1e-6
Psi_p = (Np * Rp + delta * diag(R0)) / (Np + delta)
llhs = log_F(R0, Psi_p, N, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value; setting delta = ", delta), ylab = "log-likelihood of matrix F-distribution")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))
```


```{r}
N = nrow(X_in)
Np = nrow(X_out)
nu_vec = c(1:100) 
delta = 1e-2
Psi_p = (Np * Rp + delta * diag(R0)) / (Np + delta)
llhs = log_F(R0, Psi_p, N, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value; setting delta = ", delta), ylab = "log-likelihood of matrix F-distribution")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))
```
When $\delta = 0.01$ we have larger $\nu$, which is interesting. Further increasing $\delta$ does not lead to a bigger optimal $\nu$.

```{r}
N = nrow(X_in)
Np = nrow(X_out)
nu_vec = c(1:100) 
delta = 1e-1
Psi_p = (Np * Rp + delta * diag(R0)) / (Np + delta)
llhs = log_F(R0, Psi_p, N, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value; setting delta = ", delta), ylab = "log-likelihood of matrix F-distribution")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))
```



Now let's try to add $\text{diag}(R')$ instead of $\text{diag}(R_0)$. The optimal $\nu$ is not as large. So it seems like setting the prior as $\text{diag}(R_0)$ is better.


```{r}
N = nrow(X_in)
Np = nrow(X_out)
nu_vec = c(1:100) 
delta = 1e-2
Psi_p = (Np * Rp + delta * diag(Rp)) / (Np + delta)
llhs = log_F(R0, Psi_p, N, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value; setting delta = ", delta), ylab = "log-likelihood of matrix F-distribution")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))
```


### Model 5

$$\mathrm{IW}(R_0 | \nu \Psi', \nu + p + 1)$$

```{r}

delta = 1e-6
Psi_p = (Np * Rp + delta * diag(Rp)) / (Np + delta)
llhs = log_iw(R0, Psi_p, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value; setting delta = ", delta), ylab = "log-likelihood of IW")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))

```


```{r}

delta = 1e-2
Psi_p = (Np * Rp + delta * diag(Rp)) / (Np + delta)
llhs = log_iw(R0, Psi_p, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value; setting delta = ", delta), ylab = "log-likelihood of IW")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))

```

They are both bad, because in the pdf of IW, we take the inverse of $R_0$ not $R'$ (out-of-sample cov. mat.). So adding a small regularization to $R'$ does not help and the PDF is small due to the singularity of $R_0$. I can not make $\nu$ any larger with Inverse-Wishart formulation here. But recall that in the Variational Inference (jointly with SuSiE), the estimated $R_0$ does borrow eigenvectors of $R'$ so that the estimated $\nu$ might be larger. 

In conclusion, jointly learning $\delta$ can help to achieve larger $\nu$, hence borrowing more strength from the out-of-sample LD (if appropriately adjusted by $\delta$).


==============

Extra: We try the model

$$\mathrm{F}(R_0 | \frac{\nu \Psi''}{N}, N, \nu + 2)$$
where $\Psi'' = \dfrac{N' \text{diag}(R_0)^{1/2} \text{diag}(R')^{-1/2} R' \text{diag}(R')^{-1/2} \text{diag}(R_0)^{1/2} + \delta \text{diag}(R_0)}{N' + \delta}$. But the estimated optimal $\nu$ seems not as large.


```{r}
N = nrow(X_in)
Np = nrow(X_out)
nu_vec = c(1:100) 
delta = 1e-2
Psi_pp = (Np * diag(diag(R0)^(0.5)) %*% cov2cor(Rp) %*% diag(diag(R0)^(0.5)) + delta * diag(R0)) / (Np + delta)
llhs = log_F(R0, Psi_p, N, nu_vec)
plot(nu_vec, llhs, xlab = paste0("nu value when delta = ", delta), ylab = "log-likelihood of matrix F-distribution")

print(paste0("Optimal nu is ", nu_vec[which.max(llhs)]))

```


