---
title: Examining the SuSiE residuals with an in-sample and an out-of-sample LD matrix
author: Peter Carbonetto
output: workflowr::wflow_html
---

In this exploratory analysis I'm using a slightly modified version of
susieR that can be installed in this way:

```{r install-susie, eval=FALSE}
library(remotes)
install_github("stephenslab/susieR@output_residuals")
```

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

## Initial setup

Here I set the seed for reproducibility.

```{r setup}
library(susieR)
library(matrixStats)
print(packageVersion("susieR"))
set.seed(1)
p1      <- 2
min_maf <- 0.03
n_out   <- 100
```

Load a version of the GTEx genotype data:

```{r load-genotype}
X <- readRDS("data/Thyroid.FMO2.1Mb.RDS")$X
storage.mode(X) <- "double"
X <- X[,1:1000]
```

Filter the SNPs by MAF:

```{r filter-by-maf}
maf <- colMeans(X)/2
maf <- pmin(maf,1 - maf)
j   <- which(maf > min_maf)
X   <- X[,j]
```

Dimensions of the genotype matrix we will use:

```{r dim-x}
print(dim(X))
```

## Simulate data set

Select the causal SNPs, then generate the (sparse) true effects:

```{r simulate-effects}
p        <- ncol(X)
j        <- sample(p,p1)
b        <- rep(0,p)
names(b) <- colnames(X)
b[j]     <- sample(c(-1,1),p1,replace = TRUE)
```

Now simulate y:

```{r simulate-y}
n <- nrow(X)
e <- rnorm(n,sd = 0.85)
y <- drop(X %*% b + e)
y <- y/sd(y)
```

Compute the "in-sample" LD matrix and an "out-of-sample" LD matrix
(more accurately this is an LD matrix obtained using a subset of the
samples):

```{r compute-ld}
R <- cor(X)
i <- sample(n,n_out)
R_out <- cor(X[i,])
```

Compute the association statistics (z-scores):

```{r z-scores}
out  <- univariate_regression(X,y)
zhat <- with(out,betahat/sebetahat)
```

## Run SuSiE-RSS with the true number of causal SNPs

Run SuSiE-RSS with the in-sample and out-of-sample LD matrices when
the number of CSs matches the true number of causal SNPs:

```{r susie-L2}
fit <- susie_rss(zhat,R,L = 2,n = n,estimate_prior_method = "EM",
                 tol = 1e-6,min_abs_corr = 0,verbose = TRUE)
fit_out <- susie_rss(zhat,R_out,L = 2,n = n,estimate_prior_method = "EM",
                     tol = 1e-6,min_abs_corr = 0,verbose = TRUE)
```

Both of the SuSiE-RSS analyses produce a similar result: two CSs, with
each CS recovering a causal SNPs.

```{r susie-plots-1, fig.height=2.25, fig.width=6}
par(mfrow = c(1,2),mar = c(4,4,2,1))
susie_plot(fit,y = "PIP",b = (b != 0),xlab = "SNP",main = "in-sample LD")
susie_plot(fit_out,y = "PIP",b = (b != 0),xlab = "SNP",
           main = "out-of-sample LD")
```

So this seems promising. It is also encouraging that the distributions
of the SER residuals are similar in the two analyses:

```{r compare-residuals, fig.height=2, fig.width=6}
par(mfrow = c(1,2),mar = c(4,4,2,1))
hist(fit$XtR,breaks = 64,xlab = "residuals",ylab = "",main = "in-sample LD")
hist(fit_out$XtR,breaks = 64,xlab = "residuals",ylab = "",
     main = "out-of-sample LD")
```

## Run susie_rss again when the number of causal SNPs is treated as an unknown

Now let's see what happens in the more realistic situation when we
don't know in advance what is the number of causal SNPs. Here are I'll
set the upper limit on the number of CSs to 4.

```{r susie-L4}
fit <- susie_rss(zhat,R,L = 4,n = n,estimate_prior_method = "EM",
                 tol = 1e-6,min_abs_corr = 0,verbose = TRUE)
fit_out <- susie_rss(zhat,R_out,L = 4,n = n,estimate_prior_method = "EM",
                     tol = 1e-6,min_abs_corr = 0,verbose = TRUE)
```

With the in-sample LD, the results look very similar to the previous
result—so SuSiE was able to correctly determine the number of causal
SNPs.

But the result looks much different with the out-of-sample LD: it
keeps all four CSs and fails to recover one of the causal SNPs.

```{r susie-plots-2, fig.height=2.25, fig.width=6}
par(mfrow = c(1,2),mar = c(4,4,2,1))
susie_plot(fit,y = "PIP",b = (b != 0),xlab = "SNP",main = "in-sample LD")
susie_plot(fit_out,y = "PIP",b = (b != 0),xlab = "SNP",
           main = "out-of-sample LD")
```
