---
title: Notes on the Wishart and inverse Wishart distributions
author: Peter Carbonetto
output: workflowr::wflow_html
---

In the *Bayesian data analysis* book (3rd ed), the authors make the
following claim: *Setting ${\bf\Sigma} \sim
\mbox{Inv-Wishart}_{d+1}({\bf I})$ has the appealing feature that each
of the correlations in ${\bf\Sigma}$ has, marginally, a uniform prior
distribution.*
[Note: ${\bf\Sigma}$ and ${\bf I}$ are $d \times d$ matrices.] Here I
check this claim in a few small simulations.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the packages needed for this analysis:

```{r load-pkgs}
library(ggplot2)
library(cowplot)
```

Set the seed to make the results reproducible:

```{r set-seed}
set.seed(1)
```

Here's a function I will use to simulate covariance matrices from the
Wishart or inverse Wishart with $\nu = d + 1$, and then output
(without loss of generality) the simulated correlations between
dimensions 1 and 2:

```{r sim-wishart}
sim_wishart <- function (n, d, v = d + 1, S = diag(d),
                         dist = c("wish", "invwish")) {
  X <- rWishart(n,v,S)
  for (i in 1:n) {
    x <- X[,,i]
	if (dist == "invwish")
      x <- solve(x)
    x <- cov2cor(x)
    X[,,i] <- x
  }
  return(drop(X[1,2,]))
}
```

This is a function I will use to create the histograms to visualize
the distribution of the correlations:

```{r create-cor-hist}
create_cor_hist <- function (x, title = "") {
  pdat <- data.frame(cor = x)
  return(ggplot(pdat,aes(x = x)) +
    geom_histogram(bins = 24,fill = "black",color = "black") +
	labs(x = "correlation",title = title) +
	theme_cowplot(font_size = 12) +
	theme(plot.title = element_text(size = 12,face = "plain")))
}
```

In 2 dimensions, the claim appears to hold for the inverse Wishart:

```{r invwish-2d-unif, fig.height=2.5, fig.width=2.5}
x <- sim_wishart(1000,2,dist = "invwish")
create_cor_hist(x,title = "IW(d+1,I), d=2")
```

As we increase the number of degrees of freedom, the correlations
become more concentrated around zero:

```{r invwish-2d-nonunif, fig.height=2.5, fig.width=2.5}
x <- sim_wishart(1000,2,v = 6,dist = "invwish")
create_cor_hist(x,title = "IW(d+4,I), d=2")
```

Interestingly, the Wishart distribution with $d + 1$ degrees of
freedom also generates a uniform distribution for the correlations, at
least in 2 dimensions:

```{r wish-2d-unif, fig.height=2.5, fig.width=2.5}
x <- sim_wishart(1000,2,dist = "wish")
create_cor_hist(x,title = "W(d+1,I), d=2")
```

However, for more than 2 dimensions, the distribution of correlations
is no longer uniform:

```{r wish-3d-unif, fig.height=2.5, fig.width=2.5}
x <- sim_wishart(1000,3,dist = "wish")
create_cor_hist(x,title = "W(d+1,I), d=3")
```

By contrast, the inverse Wishart retains a uniform distribution:

```{r invwish-3d-unif, fig.height=2.5, fig.width=2.5}
x <- sim_wishart(1000,3,dist = "invwish")
create_cor_hist(x,title = "IW(d+1,I), d=3")
```

This illustrates the ability of the inverse Wishart to maintain a
"diffuse" prior in higher dimensions.
